Paper

### 2017-11-17 Paper.Rmd
...

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))

# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

@@

nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))

globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
#globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

@@

pt.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Patient = pt.status))

@@

ptbe.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))

genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status )) + labs(title='NDBE samples only')

@@

cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status ), type='bar') + labs(title='NDBE samples only')
```

## Individual samples
```{r}

a = sample(subset(patient.info, Status == 'NP' & Pathology %in% c('BE'))$Samplename, 1)
b = sample(subset(patient.info, Status == 'P' & Pathology %in% c('BE'))$Samplename, 1)

cn.mtn.plot(subset(globalCNdf, variable %in% c(a) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))
cn.mtn.plot(subset(globalCNdf, variable %in% c(b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))

cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples")
cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual NDBE samples")

@@

```{r echo=F, fig.height=5, fig.width=6}
globalCNdf$g.start = globalCNdf$start
for (i in 2:nrow(chr.lengths)) {
  rows = which(globalCNdf$chr == chr.lengths[i, 'chrom'])
  globalCNdf[rows, 'g.start'] = globalCNdf[rows, 'start'] + chr.lengths[i-1, 'genome.length']
}

p1 = ggplot(globalCNdf) + geom_tile(aes(x=g.start, y=variable, fill=value, group=Status)) +
  scale_fill_gradient2(high='darkgreen', low='darkred', mid='white') +
  labs(x='', y='Sample') + scale_x_continuous(expand=c(0,0)) +
  #scale_x_continuous(expand=c(0,0), breaks=chr.lengths$genome.length, labels=chr.lengths$chrom) +
  geom_vline(data=chr.lengths, aes(xintercept=genome.length), color='grey') +
  plot.theme + theme(axis.ticks=element_blank(), axis.text=element_blank(),panel.grid.major=element_blank(), legend.position='bottom')

p1leg = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')

p2 = ggplot(globalCNdf, aes(y=variable)) + geom_tile(aes(x=0, fill=Status), show.legend = F) + 
  labs(x=' ', y='') + scale_fill_manual(values=c('darkblue','darkred')) + scale_x_continuous(expand=c(0,0)) +
  plot.theme + theme( axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), panel.grid=element_blank())

grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(10,1), heights=c(10,1))

@@

get.roc.stat<-function(roc) {
  cbind.data.frame('AUC'=pROC::auc(roc), 
        'Specificity'=round(pROC::coords(roc, 'best')[['specificity']], 2), 
        'Sensitivity'=round(pROC::coords(roc, 'best')[['sensitivity']], 2),
        'model'=roc$model)
}

m = rbind.data.frame(get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))

m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)

@@

aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model)))

df = do.call(rbind,
             lapply(rocList, function(r) 
              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),
                               'model'=r$model)))
  ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) + geom_line(show.legend = F) + scale_x_reverse() + 
    geom_label(data=m, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC*100), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
  labs(title='ROC Curves', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

ggplot(melt(m,id.vars = 'model'), aes(model, value*100, group=variable, fill=variable)) + ylim(0,100) +
  geom_col(position="dodge") + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep='') ), position=position_dodge(width=1), vjust=-0.5) + 
  labs(x='', y='Percent', title='ROC') +
  scale_fill_manual(values=c('dodgerblue3','firebrick4', 'green4'), name='') + plot.theme + 
  theme(legend.position = 'bottom', axis.text.x=element_text(angle=45,hjust=1))

@@

@@

  } else {
    return( cbind('HGD'=pred(h), 'total.h'=nrow(h), 
                  'LGD'=pred(l), 'total.l'=nrow(l), 
                  'ID'=pred(id), 'total.id'=nrow(id),
                  'NDBE'=pred(be), 'total.be'=nrow(be)) )
    }
}

collapse.pred.by.path<-function(pts, p53=F, predFUN = NULL) {
  if (!is.null(predFUN))
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53, predFUN) )
  else
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53) )

@@

predP53 = function(rows) length(which(rows$p53.Status == 1 ))
p53byPath = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], T, predP53)

@@

m = rbind(global.melt, np.melt, p53.melt)

@@

m$pred = factor(m$pred)
m$final = ifelse(m$path == 'HGD', T, F)

ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
  geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('firebrick4','green4','slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed','p53 IHC'))  +
  geom_text(aes(vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')),
            position=position_stack(),color='white',show.legend=F) +
  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
  geom_text(data=p53.melt, aes(path,value,vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
  labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom')

@@

@@

```{r, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
transform.pt.by.time<-function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
                  'pred'=sum(as.integer(Prediction > cutoff)), 'samples'=length(PID), 
                  #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12, 
                  'max.path'=max(Pathology))
  arrange(pt, -Endoscopy.Year, PID)
}

back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], transform.pt.by.time)

endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
  }))
  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient

  df$Patient = rownames(df)
  return(df)
}))

endos = endos[complete.cases(endos),]
endos$i = endos$i-1

endos$pred.ratio = round(endos$pred/endos$samples, 2)
endos = transform(endos, Patient=reorder(Patient, -months.before.final) )

endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12))
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

@@

## Flip order, 0-1 comes last
m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)), 'n.samp'=sum(samples) )

m$highlight = F
m[4:6,'highlight'] = T
m$final = ifelse(m$brks == '[0,1]', T, F)

ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='dodgerblue3', show.legend = F) +     
  scale_fill_manual(values=c('dodgerblue3', 'yellowgreen')) +
  geom_text(aes(y=pred.r*100+1.5, label=paste(round(pred.r*100), '%', sep=''))) +
  geom_text(aes(y=pred.r*100-2, label=paste('n=',n,'\ns=',n.samp,sep='')), color='white') + 
  scale_x_discrete(labels=m$labels) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))

@@

endos$months.before.final = endos$months.before.final*-1
me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

@@

```{r, spacial, fig.width=8, fig.height=6}
transform.by.time<-function(pt) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
  #'nl' = Block
  #'nl' = str_extract(Path.ID, "B\\w?[0-9]+")
  )

  pt = merge(pt, subset(patient.info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename')

  pt$nl = (pt$Block-1)*2
  #pt$nl = (as.numeric(factor(pt$nl, levels=(unique(sort(pt$nl))), ordered = T))-1)*2
  #table(pt[,c('nl','months.before.final')])

>>>>

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

@@

nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
#globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

@@

pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))

@@

ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')

@@

#cn.mtn.plot(dfpts,  labeller(Status = ptbe.status ), type='bar') + labs(title='NDBE samples only')

a = sample(subset(patient.info, Status == 'NP' )$Samplename, 1)
b = sample(subset(patient.info, Status == 'P' )$Samplename, 1)

cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual samples")
cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual samples")

@@

```{r echo=F, fig.height=5, fig.width=6}
globalCNdf$g.start = globalCNdf$start
globalCNdf$g.end = globalCNdf$end
for (i in 2:nrow(chr.lengths)) {
  rows = which(globalCNdf$chr == chr.lengths[i, 'chrom'])
  globalCNdf[rows, 'g.start'] = globalCNdf[rows, 'start'] + chr.lengths[i-1, 'genome.length']
  globalCNdf[rows, 'g.end'] = globalCNdf[rows, 'end'] + chr.lengths[i-1, 'genome.length']
}

globalCNdf = arrange(globalCNdf, chr, start)
globalCNdf$segment = with(globalCNdf, paste(chr, ':', start, '-', end, sep=''))

globalCNdf$chr = factor(globalCNdf$chr, levels=c(1:22), ordered = T)
globalCNdf$Status = factor(globalCNdf$Status)

p1 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(segment, variable, fill=value, group=Status)) + geom_tile() +
  scale_fill_gradient2(high='red3', low='green4', mid='white')  + 
  facet_grid(~chr, scales = 'free_x', space = 'free') + plot.theme + labs(x='',y='Samples') + 
  theme(axis.ticks=element_blank(), axis.text=element_blank(), panel.spacing.x = unit(0, 'lines'), legend.position='bottom')

p1leg = get.legend(p1)
p1 = p1 + theme(legend.position = 'none')

p2 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(y=variable, group=Status)) + 
  geom_tile(aes(x=0, fill=Status), show.legend = F) + 
  labs(x=' ', y='') + scale_fill_manual(values=c('darkblue','darkred')) + scale_x_continuous(expand=c(0,0)) +
  plot.theme + theme( axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), panel.grid=element_blank())

grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(20,1), heights=c(10,1))

@@

get.roc.stat<-function(roc) {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

  sens = c(pROC::coords(roc, 'best')[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, 'best')[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, 'best')[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, 'best')[['sensitivity']], conf.level=0.95))

  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
  colnames(stat) = c ('value','CI.min','CI.max')
  stat = as.data.frame(stat)
  stat$model = roc$model
  stat$Measure = rownames(stat)
  stat
}

m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)

@@

aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))

df = do.call(rbind, lapply(rocList, function(r) 
              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model) 
              ))

ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) + geom_line(show.legend = F) + scale_x_reverse() + 
    geom_label_repel(data=aucs, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
  labs(title='ROC Curves', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

dodge = position_dodge(width=0.9)
ggplot(melt(m, measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
  theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison')

@@

```{r, spacial, fig.width=8, fig.height=6}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
  )

  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = table(pt$Endoscopy.Year, pt$PID)
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
      tmp = pt[which(pt$Endoscopy.Year == yr),]
      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))

      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
  }

  pt$nl = (pt$Block-1)*2

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status')])

cuts = seq(0,1,0.1)
#cuts = quantile(back$Prediction, probs=seq(0,1,1/5))

back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

ft.fun<-function(NP,P) {
 f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
 cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

riskCols = c('red3','orange','green')
ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability')

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')

ggplot(back, aes(1:nrow(back), Prediction)) + geom_point(aes(color=Risk, shape=Status)) + labs(x='sample', y='model probability', title='Per-sample probabilities')

back = as.data.frame(back %>% group_by(Risk) %>% mutate( 'mean(P)'=mean(Prediction), 'se(P)'=sd(Prediction)/sqrt(length(Prediction)), 'sd(P)'=sd(Prediction) ))
```

At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile. Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges. This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.

`r pander(as.data.frame(pred.confidence))`

@@

  } else {
    return( cbind('HGD'=pred(h), 'total.h'=nrow(h), 
                  'LGD'=pred(l), 'total.l'=nrow(l), 
                  'ID'=pred(id), 'total.id'=nrow(id),
                  'NDBE'=pred(be), 'total.be'=nrow(be)) )
    }
}
collapse.pred.by.path<-function(pts, p53=F, predFUN = NULL) {
  if (!is.null(predFUN))
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53, predFUN) )
  else
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53) )

@@

predP53 = function(rows) length(which(rows$p53.Status == 1 ))
p53byPath = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], T, predP53)
p53.melt = melt(p53byPath, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
p53.melt$Status = ifelse(p53.melt$variable == 'prog.ratio', 'Progressor','Non-Progressor')

@@

path.m = rbind(global.melt, np.melt, p53.melt)

@@

path.m$pred = factor(path.m$pred)
path.m$final = ifelse(path.m$path == 'HGD', T, F)

ggplot(path.m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
  geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('firebrick4','green4','slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed','p53 IHC'))  +
  geom_text(aes(vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')),
            position=position_stack(),color='white',show.legend=F) +
  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
  geom_text(data=p53.melt, aes(path,value,vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
  labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom')

@@

Same plot, but mapped to the risk categories instead.

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
back$Pathology = as.character(back$Pathology)
back$Pathology[back$Pathology %in% c('HGD','IMC')] = 'HGD/IMC'
back$Pathology = factor(back$Pathology, levels=c('BE', 'ID','LGD','HGD/IMC'), ordered = T)

riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
riskPath$Risk = factor(riskPath$Risk, levels=c('High','Moderate','Low'), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))

ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('red','orange2','green3')) + 
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
```

@@

```{r, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
endos = subset(back, Status == 'P') %>% group_by(months.before.final, PID) %>% summarise(
              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
              'n.endos'=length(unique(PID)), 
              'n.samples'=length(PID),
              'pred.samples'=length(which(Risk == 'High')),
              'max.path'=max(Pathology))

endos %>% group_by(months.before.final) %>% summarise(sum(pred.endo)/sum(n.endos))
endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12))
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

@@

endos.m = endos %>% group_by(brks, labels) %>% summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),
  'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

rev(x$brks)

 subset(back, Status == 'P') %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction), 'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

se = unique(subset(back, Risk == 'High')$'se(P)')

ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-se)*100, ymax=(pred.r+se)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
  geom_errorbar(color='grey21') +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  scale_x_discrete(labels=endos.m$labels) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))

## Flip order, 0-1 comes last
se = diff(range(subset(pred.confidence, Risk == 'High')$perc))
endos.m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks),'n.success.endo'=length(which(pred>0)), 'n.endo'=length((Patient)), 
                                                          'n.successes.samp'=sum(pred), 'n.samp'=sum(samples)) %>% mutate( 
                                                          'CI.min'=binom::binom.confint(n.success.endo, n.endo, conf.level=0.95, methods='asymptotic')[['lower']],
                                                          'CI.max'=binom::binom.confint(n.success.endo, n.endo, conf.level=0.95, methods='asymptotic')[['upper']])
ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=CI.min*100, ymax=ifelse(CI.max>1, 100, CI.max*100), group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + geom_errorbar(width=0.25, color='grey21') +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  scale_x_discrete(labels=endos.m$labels) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))

endos.samp = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'=sum(pred)/sum(samples), 'n.successes.samp'=sum(pred), 'n.samp'=sum(samples)) %>% mutate( 
                                                          'CI.min'=binom::binom.confint(n.successes.samp, n.samp, conf.level=0.95, methods='asymptotic')[['lower']],
                                                          'CI.max'=binom::binom.confint(n.successes.samp, n.samp, conf.level=0.95, methods='asymptotic')[['upper']])
ggplot(endos.samp, aes(x=brks, y=pred.r*100, ymin=CI.min*100, ymax=ifelse(CI.max>1, 100, CI.max*100), group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='darkred', show.legend = F) + geom_errorbar(width=0.25, color='grey21') +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('s=',n.samp,sep='')), color='white') + 
  scale_x_discrete(labels=endos.m$labels) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Sample predictions prior to final endoscopy', subtitle='s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))

@@

endos$months.before.final = endos$months.before.final*-1
me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))

m = back %>% group_by (Status, brks, Risk) %>% summarise( 'count'=length(Risk) ) %>% mutate('total.by.brk'=sum(count)) %>% mutate('ratio'=round(count/total.by.brk,3)*100)
m$Status = ifelse(m$Status == 'P', 'Progressor','Non-Progressor')
ggplot(m, aes(brks, ratio)) + geom_bar(stat='identity', aes(group=Risk, fill=Risk)) + 
  geom_text(aes(group=Risk, label=paste(signif(ratio, 2),'%',sep='')), position = position_stack(vjust = 0.5)) +
  geom_text(data=(m %>% group_by(Status, brks) %>% summarise('total'=unique(total.by.brk))), aes(label=paste('n=',total,sep=''), y=101)) +
  scale_fill_manual(values=c('red','orange2','green3')) + 
  facet_grid(~Status) + plot.theme + labs(x='Months prior to endpoint', y='Percent predicted',title='Predictions prior to final sample')

@@

`r pander(as.data.frame(pred.confidence))`

The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".

```{r, fig.width=5, fig.height=10 }
pred.tiles <- function(df, probabilities=F) {
  p = ggplot(df, aes(brks, nl))
